Postural developments mediate children's visual access to social information

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.Rmd
detector wording

Here, we directly ask whether postural and locomotive developments change the availability of social information--the presence of faces and hands. To do so, we recorded the visual experience of a group of infants in three age ranges (8,12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated. We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (operationalized by the presence of wrists) in the child’s visual environment, both overall and relative to naming events by their caregivers. We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, and walking toddlers would have access to a richer visual landscape that would contain a larger portion of the social information in the environment.

@@

Our final sample consisted of 36 infants and children, with 12 participants in three age groups: 8 months (6 females), 12 months (7 females), and 16 months (6 females). Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80\% English at home. Demographics and exclusion rates are given in Table \ref{tab:pop}. 
\begin{table}[H]
\centering
\begin{tabular}{rrrrrr}
  \hline
 Group & N & \% incl. & Mean age & Videos length (min) \\ 
  \hline
   8 mo. &   12 & 0.46 & 8.71 & 14.41 \\ 
   12 mo. &  12 & 0.40 & 12.62 & 13.48 \\ 
   16 mo. &  12 & 0.31 & 16.29 & 15.00\\ 
   \hline
\end{tabular}
\caption{\label{tab:pop} Demographics by age group.}
\end{table}
To obtain this final sample, we tested `r length(all_subs$reason)` children, excluding  `r sum(all_subs$reason!="included")` children for the following reasons:
`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
`r sum(all_subs$reason=="would not wear camera")` for failing to wear the headcam,
`r sum(all_subs$reason=="less than 4 min HC footage")` for fewer than 4 minutes of headcam footage,
`r sum(all_subs$reason=="multiple people")` for having multiple adults present,
`r sum(all_subs$reason=="no CDI")` for missing Communicative Development Inventory (CDI) data,
`r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
`r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses. Some of these data were also used in @frank2013.

@@

## Face and Wrist Detection
We used three face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame. The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones. We used this as a benchmark for performance, and while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual]. We next capitalized on recent improvements in computer vision, testing the performance of two face detectors that both made use of Convolutional Neural Networks (CNNs) to extract face information. The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts belonging to an agent.

This second of the CNN-based algorithms - the pose detector - provided us with the location of an agent's wrists, which we used as a good proxy for the presence of a hand. We do so first because we did not capture the entire visual field experienced by the infants/children. Thus, even though the video may only show the presence of a wrist, the caregiver's hand may be within the child's field of view. Second, we do so because hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction. For example, if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy).

@@

The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc.) called OpenPose [@cao2017realtime; @simon2017hand; @wei2016cpm] avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose. The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose. For the purposes of our investigation we only made use of the body parts relevant to the face and hands (nose and wrists). We operationalized face detections as any frames containing a nose, as any half of a face necessarily contains a nose. We operationalized wrist detections as any frames containing either the left or the right wrist.

@@

For face detection, MTCNN outperformed OpenPose when looking at the composite F-score (`r round(perfFacesOverall$f[1],2)` vs. `r round(perfFacesOverall$f[2],2)`); thus the face presence analyses to follow all result from MTCNN detections. We found that although MTCNN and OpenPose perform comparably with the random sample, MTCNN performs better on the high density sample evaluation (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN. ViolaJones performed quite poorly relative to the other detectors, finding much less faces in the random sample.

For wrist detection, OpenPose performed moderately well (f = `r round(perfWristsOverall$f,2)`) with relatively high precision, but low recall on the randomly sampled set of frames (see Table 2). We thus analyze wrist detections, with the caveat that we are likely underestimating the proportion of wrists overall in the datatset.

>>>>

Here, we directly ask whether postural and locomotive developments change the availability of social information--the presence of faces and hands. To do so, we recorded the visual experience of a group of infants in three age ranges (8,12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated. We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrist presence as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers. We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, and walking toddlers would have access to a richer visual landscape that would contain a larger portion of the social information in the environment.

@@

Our final sample consisted of 36 infants and children, with 12 participants in three age groups: 8 months (6 females), 12 months (7 females), and 16 months (6 females). Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80\% English at home. Demographics and exclusion rates are given in Table \ref{tab:pop}. 
\begin{table}[H]
\centering
\begin{tabular}{rrrrrr}
  \hline
 Group & N & \% incl. & Avg age & Avg video length (min) \\ 
  \hline
   8 mo. &   12 & 0.46 & 8.71 & 14.41 \\ 
   12 mo. &  12 & 0.40 & 12.62 & 13.48 \\ 
   16 mo. &  12 & 0.31 & 16.29 & 15.00\\ 
   \hline
\end{tabular}
\caption{\label{tab:pop} Demographics by age group.}
\end{table}
To obtain this final sample, we tested `r length(all_subs$reason)` children, excluding  `r sum(all_subs$reason!="included")` children for the following reasons:
`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
`r sum(all_subs$reason=="would not wear camera")` for failing to wear the headcam,
`r sum(all_subs$reason=="less than 4 min HC footage")` for fewer than 4 minutes of headcam footage,
`r sum(all_subs$reason=="multiple people")` for having multiple adults present,
`r sum(all_subs$reason=="no CDI")` for missing Communicative Development Inventory (CDI) data,
`r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
`r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses. Some of these data were also used in @frank2013.

@@

## Face and Hand Detection
We used three face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame. The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones. We used this as a benchmark for performance, and while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual]. We next capitalized on recent improvements in computer vision, testing the performance of two face detectors that both made use of Convolutional Neural Networks (CNNs) to extract face information. The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts belonging to an agent.

This second of the CNN-based algorithms--the pose detector--provided us (along with other parts) with the location of an agent's wrists, which we used as a good proxy for the presence of a hand. We did so first because we did not capture the entire visual field experienced by the infants/children. Thus, even though the video may only show the presence of a wrist, the caregiver's hand may be within the child's field of view. Second, we did so because hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction. For example, if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy). Similarly, we used the nose (one of the 18 body parts detected) to operationalize face presence, as any half of a face necessarily contains a nose.

@@

The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc.) called OpenPose [@cao2017realtime; @simon2017hand; @wei2016cpm] avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose. The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose. For the purposes of our investigation we only made use of the body parts relevant to the face and hands (nose and wrists). We operationalized face detections as any frames containing a nose, and hand detections as any frames containing either the left or the right wrist.

@@

For face detection, MTCNN outperformed OpenPose when taking into account only the composite F-score (`r round(perfFacesOverall$f[1],2)` MTCNN vs. `r round(perfFacesOverall$f[2],2)` OpenPose); thus the face presence analyses to follow all result from MTCNN detections. We found that although MTCNN and OpenPose perform comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN. ViolaJones performed quite poorly relative to the other detectors especially on the random sample.

For wrist detection, OpenPose performed moderately well (F = `r round(perfWristsOverall$f,2)`) with relatively high precision, but low recall on the randomly sampled set of frames (see Table 2). We thus analyze wrist detections, with the caveat that we are likely underestimating the proportion of wrists overall in the datatset.

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.tex
detector wording

\begin{abstract}
The ability to process social cues--including eye gaze--is a critical
component of children's early language and cognitive development.
However, as children reach their first birthday, they begin to locomote
themselves, walking and exploring their visual environment in an
entirely new way. How do these postural and locomotive changes affect
children's access to the social information relevant for word-learning?
Here, we explore this question by using head-mounted cameras to record
infants' (8-16 months of age) egocentric visual perspective and use
computer vision algorithms to detect the proportion of faces and hands
in infants' environments. We find that infants' posture and orientation
to their caregiver largely mediate infants' access to socially
informative features, suggesting that postural and locomotive
development play a significant role in mediating children's emerging
linguistic and social capacities. Broadly, we suggest that the combined
use of head-mounted cameras and the application of new computer vision
techniques is a promising avenue for understanding the statistics of
infants' visual and linguistic experience.

@@

Here, we directly ask whether postural and locomotive developments
change the availability of social information--the presence of faces and
hands. To do so, we recorded the visual experience of a group of infants
in three age ranges (8,12, and 16 months) using head-mounted cameras
during a brief laboratory free-play session; children's posture and
orientation relative to their caregiver were also recorded from a
third-person perspective and hand-annotated. We then capitalize on
recent improvements in face and pose detection algorithms (Cao, Simon,
Wei, \& Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
frequencies of faces and hands (operationalized by the presence of
wrists) in the child's visual environment, both overall and relative to
naming events by their caregivers. We hypothesized that there would be
differential access to social information based on children's postural
developments: crawling infants would see fewer faces because they would
primarily be looking at the ground, and walking toddlers would have
access to a richer visual landscape that would contain a larger portion
of the social information in the environment.

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrrrr}
  \hline
 Group & N & \% incl. & Mean age & Videos length (min) \\ 
  \hline
   8 mo. &   12 & 0.46 & 8.71 & 14.41 \\ 
   12 mo. &  12 & 0.40 & 12.62 & 13.48 \\ 
   16 mo. &  12 & 0.31 & 16.29 & 15.00\\ 
   \hline
\end{tabular}
\caption{\label{tab:pop} Demographics by age group.}
\end{table}

@@

\subsection{Face and Wrist Detection}\label{face-and-wrist-detection}

@@

This second of the CNN-based algorithms - the pose detector - provided
us with the location of an agent's wrists, which we used as a good proxy
for the presence of a hand. We do so first because we did not capture
the entire visual field experienced by the infants/children. Thus, even
though the video may only show the presence of a wrist, the caregiver's
hand may be within the child's field of view. Second, we do so because
hands are often occluded by objects when caregivers are interacting with
children, yet still visually accessible by the child and part of their
joint interaction. For example, if a caregiver was holding a toy and
presenting it to the infant, their wrists may be visible but their hand
may not necessarily be (as it is occluded by the toy).

@@

The third algorithm was a CNN-based pose detector that provided the
locations of 18 body parts (ears, nose, wrists, etc.) called OpenPose
(Cao et al., 2017; Simon, Joo, Matthews, \& Sheikh, 2017; Wei,
Ramakrishna, Kanade, \& Sheikh, 2016) avaliable at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose. For the purposes of
our investigation we only made use of the body parts relevant to the
face and hands (nose and wrists). We operationalized face detections as
any frames containing a nose, as any half of a face necessarily contains
a nose. We operationalized wrist detections as any frames containing
either the left or the right wrist.

>>>>

\begin{abstract}
The ability to process social cues--including eye gaze--is a critical
component of children's early language and cognitive development.
However, as children reach their first birthday, they begin to locomote
themselves, walking and exploring their visual environment in an
entirely new way. How do these postural and locomotive changes affect
children's access to the social information relevant for word-learning?
Here, we explore this question by using head-mounted cameras to record
infants' (8-16 months of age) egocentric visual perspective and use
computer vision algorithms to detect the proportion of faces and hands
in infants' environments. We find that infants' posture and orientation
to their caregiver largely mediate infants' access to socially
informative features, suggesting that postural and locomotive
developments play a significant role in the emergence of children's
linguistic and social capacities. Broadly, we suggest that the combined
use of head-mounted cameras and the application of new computer vision
techniques is a promising avenue for understanding the statistics of
infants' visual and linguistic experience.

@@

Here, we directly ask whether postural and locomotive developments
change the availability of social information--the presence of faces and
hands. To do so, we recorded the visual experience of a group of infants
in three age ranges (8,12, and 16 months) using head-mounted cameras
during a brief laboratory free-play session; children's posture and
orientation relative to their caregiver were also recorded from a
third-person perspective and hand-annotated. We then capitalize on
recent improvements in face and pose detection algorithms (Cao, Simon,
Wei, \& Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
frequencies of faces and hands (using wrist presence as a proxy for the
latter) in the child's visual environment, both overall and relative to
naming events by their caregivers. We hypothesized that there would be
differential access to social information based on children's postural
developments: crawling infants would see fewer faces because they would
primarily be looking at the ground, and walking toddlers would have
access to a richer visual landscape that would contain a larger portion
of the social information in the environment.

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrrrr}
  \hline
 Group & N & \% incl. & Avg age & Avg video length (min) \\ 
  \hline
   8 mo. &   12 & 0.46 & 8.71 & 14.41 \\ 
   12 mo. &  12 & 0.40 & 12.62 & 13.48 \\ 
   16 mo. &  12 & 0.31 & 16.29 & 15.00\\ 
   \hline
\end{tabular}
\caption{\label{tab:pop} Demographics by age group.}
\end{table}

@@

\subsection{Face and Hand Detection}\label{face-and-hand-detection}

@@

This second of the CNN-based algorithms--the pose detector--provided us
(along with other parts) with the location of an agent's wrists, which
we used as a good proxy for the presence of a hand. We did so first
because we did not capture the entire visual field experienced by the
infants/children. Thus, even though the video may only show the presence
of a wrist, the caregiver's hand may be within the child's field of
view. Second, we did so because hands are often occluded by objects when
caregivers are interacting with children, yet still visually accessible
by the child and part of their joint interaction. For example, if a
caregiver was holding a toy and presenting it to the infant, their
wrists may be visible but their hand may not necessarily be (as it is
occluded by the toy). Similarly, we used the nose (one of the 18 parts)
to operationalize face presence, as any half of a face necessarily
contains a nose.

@@

The third algorithm was a CNN-based pose detector that provided the
locations of 18 body parts (ears, nose, wrists, etc.) called OpenPose
(Cao et al., 2017; Simon, Joo, Matthews, \& Sheikh, 2017; Wei,
Ramakrishna, Kanade, \& Sheikh, 2016) avaliable at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose. For the purposes of
our investigation we only made use of the body parts relevant to the
face and hands (nose and wrists). We operationalized face detections as
any frames containing a nose, and hand detections as any frames
containing either the left or the right wrist.
